Does stomatal patterning in amphistomatous leaves minimize the CO\(_2\) diffusion path length within leaves?

### 2024-03-04 ms/response-to-reviewers.qmd

There are several descriptions used that are not entirely clear. What exactly is meant by "over dispersal" in the context of stomatal patterning. Does this mean there are too many stomata on the leaf or their spacing is not optimal but the correct number? This is not common terminology when referring to stomatal patterning. The assessment of distribution relative to a uniform random distribution needs further explanation as it is well established that stomata are randomly distributed.

**Response:** \textcolor{blue}{We appreciate the reviewer's time to read and comment on our manuscript. They have brought up important points of clarification that must be addressed before the manuscript is ready to be published.}

\textcolor{blue}{We have clarified and standardized our terminology, adding explanations when needed. It has long been known that stomata are non-randomly distributed, tending to be more uniformly distribution than random (see Croxdale et al. (2000) for a nice review of historical evidence of stomatal patterning). In Croxdale et al. (2000), there is also confusion over terminology in the published literature, so we are not sure that there are field-accepted terms. So, we have adopted our own set of terminology which mirrors that of the field and related fields which discuss the spatial patterns of discrete objects. Specifically, see lines 111-117:}

> "We compared observed stomatal patterning to an ideal pattern (uniform equilateral triangular grid) and a null model (random uniform distribution). The terminology is unfortunately confusing because the word `uniform' is used in different ways. A uniform equilateral triangular grid means that the distance between stomata is uniform; a random uniform distribution means that a stomate has an equal probability (i.e. uniform) of occurring anywhere on the leaf surface. To limit confusion, we refer to the ideal pattern (equilateral triangle grid) as uniform and the null pattern (random uniform) as random. When observed stomatal patterns are more dispersed than expected under random patterning, we refer to this as overdispersed. Note, however, that overdispersed compared to random is still less dispersed than ideal, because the ideal pattern is maximally dispersed."

### 2024-03-04 ms/ms.Rmd

We compared observed stomatal patterning to an ideal pattern (uniform equilateral triangular grid) and a null model (random uniform distribution). The terminology is unfortunately confusing because the word `uniform' is used in different ways. A uniform equilateral triangular grid means that the distance between stomata is uniform; a random uniform distribution means that a stomate has an equal probability (i.e. uniform) of occurring anyone on the leaf surface. To limit confusion, we refer to the ideal pattern as uniform (equilateral triangle grid) and the null pattern as random (random uniform). When observed stomatal patterns are more dispersed than expected under random patterning, we refer to this as overdispersed. Note, however, that overdispersed compared to random is still less dispersed than ideal, because the ideal pattern is maximally dispersed.

>>>>

We compared observed stomatal patterning to an ideal pattern (uniform equilateral triangular grid) and a null model (random uniform distribution). The terminology is unfortunately confusing because the word `uniform' is used in different ways. A uniform equilateral triangular grid means that the distance between stomata is uniform; a random uniform distribution means that a stomate has an equal probability (i.e. uniform) of occurring anywhere on the leaf surface. To limit confusion, we refer to the ideal pattern (equilateral triangle grid) as uniform and the null pattern (random uniform) as random. When observed stomatal patterns are more dispersed than expected under random patterning, we refer to this as overdispersed. Note, however, that overdispersed compared to random is still less dispersed than ideal, because the ideal pattern is maximally dispersed.

### 2024-03-04 ms/ms.tex

We compared observed stomatal patterning to an ideal pattern (uniform
equilateral triangular grid) and a null model (random uniform
distribution). The terminology is unfortunately confusing because the
word `uniform' is used in different ways. A uniform equilateral
triangular grid means that the distance between stomata is uniform; a
random uniform distribution means that a stomate has an equal
probability (i.e.~uniform) of occurring anyone on the leaf surface. To
limit confusion, we refer to the ideal pattern as uniform (equilateral
triangle grid) and the null pattern as random (random uniform). When
observed stomatal patterns are more dispersed than expected under random
patterning, we refer to this as overdispersed. Note, however, that
overdispersed compared to random is still less dispersed than ideal,
because the ideal pattern is maximally dispersed.

>>>>

We compared observed stomatal patterning to an ideal pattern (uniform
equilateral triangular grid) and a null model (random uniform
distribution). The terminology is unfortunately confusing because the
word `uniform' is used in different ways. A uniform equilateral
triangular grid means that the distance between stomata is uniform; a
random uniform distribution means that a stomate has an equal
probability (i.e.~uniform) of occurring anywhere on the leaf surface. To
limit confusion, we refer to the ideal pattern (equilateral triangle
grid) as uniform and the null pattern (random uniform) as random. When
observed stomatal patterns are more dispersed than expected under random
patterning, we refer to this as overdispersed. Note, however, that
overdispersed compared to random is still less dispersed than ideal,
because the ideal pattern is maximally dispersed.
